\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-13 arch.tex
updates

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at any single tick \comment{at
  every single tick??? maybe} to ensure that all reactors are aware of
the current state of the world. \texttt{step} relates to a single
decision step of the planning process which is focused on the future
evolution of state variables; it can take multiple steps spanning
multiple ticks for a reactor to produce a plan that is both complete
and valid. When such plan is identified \comment{how? Is there a
  terminating condition?} its {\em external} timeline(s) are then
dispatched to the corresponding reactors which own these times and
which in turn can start their own deliberation \texttt{step}s to
satisfy these new objectives. Consequently execution occurs as these
goals and plans propagate down in the reactor hierarchy until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal into a command for
execution in hardware.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

The manner in which the agent interleaves planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization occurs at every tick
interrupting planning; this allows a reactor to identify its current
state that can propagate through the reactor hierarchy in order to
ensure a consistent view of the state of the world between all the
reactors.
% This view of a reactor has two parallel tasks allowing for long
% planning deliberation while ensuring that the state evolution
% propagates through the agent as time advance.
These intertwined tasks together provide two natural information flow
methods (bottom-up for state estimation and top-down for plan
projection) that we expect to see in a control loop. As they are both
inference-based processes it is natural to implement a reactor based
on generic automated planning. For example in Fig. \ref{fig:agent},
apart from \textsf{Shore comm.} and \textsf{Vehicle} acting as
interfaces to the system, every other reactor (in blue) are different
instances of a \eu reactor differentiated only by their model and \eu
solver configurations.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is
to deliberate as time advance. The design of this reactor takes cues
from the IDEA architecture presented on \cite{mus02, mus06} in the
sense that both planning and execution are based on a single unified
model. The choice of \eu as the planning framework was directed by the
fact that this framework provide a rich and highly configurable
toolset to be able to infer around a rich temporal domain.

This reactor is implemented following the idea that in \rx, the
whole execution cycle is two deliberation processes. Both of these
processes are implemented as two europa solvers modifying the same
plan database and for which the execution is managed by the \rx 
framework as time advance during the execution of the system:

@@

For the sake of simplicity we introduce first how a reactor can track
and identify its state in the context where it does not have any
compelling need to deliberate. Assume here that we have a reactor that
have no future goal. {\em Need to develop the reason why it is a
  necessity: primarily to provide its internal state to whoever
  observe it but als simply to ensure that its current representation
  of the world is up to date and still consistent}

In the europa based reactor we introduce this requirement as a new
type of flaw for the europa framework. This flaw enforces at the
reactor to identify fully its internal state for the current tick. By
using this new flaw we can describe the synchronization process as
this general sequence :

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

The resolution of internal state flaw can be resolved using one of the
following choices which are evaluated in the given order:

@@

\item Extend the previous state value to end after this tick (ie
  restrict its end time to $[\tau+1, \infty)$).

\item Start the next active token in the timeline and
  attempt to start it now (ie restrict its start time to the single
  value $\tau$).

@@

All of these choices are evaluated sequentially until a consistent
solution with no more flaw for this tick is identified. In order to do
so we need to make the assumption that the current state value of an
internal timeline do not depend on the future or more accurately that
any choices made during this synchronization will not lead to a future
inconsistency (meaning no possible solution) for a future
synchronization. Such assumption deeply impact the set of possible
domains one the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model :
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   equals(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11 }
\end{verbatim}
While this model is perfectly standard and acceptable in the overall
case. It can become problematic while tied to the
execution. Specificially it puts a strong tie between the {\em fact} that we
are \texttt{Going} to a location and the {\em expected} future outcome  of
ending at this location. This ties is even stronger due to the
constraint at line 5. Knwoing this aspect assume that the
\texttt{Agent.location} is {\em external} to the reactor which receive
the observation \texttt{Going(Home,SuperMarket)} which synchronization
resolved by producing the observation \texttt{Go(Home, SuperMarket)}
in the Agent timeline. From this point this means that the only
possible outcome on \texttt{Agent.location} is to be
\texttt{At(SuperMarket)} in the directly foreseeable future. In the
real world the fact that we are \texttt{Going} to a certain location
does not necessarly imply that we will succeed on ending up there. An
extreme case could be that due to some exogenous event, the road to
the \texttt{SuperMarket} is closed which could result on an alternate
observation \texttt{At(7th st. \& Foo Ave.)}. This observation
contradicts thre existing plan the reactor maintin and it is important
here to relaize the temp[oral direction of the causal link that is
triggereing this inconsitency. Indeed at the time we will observe this
contradicting \texttt{At}, the rule that generated this conlict is
connected to the past observations \texttt{Going} and \texttt{Go}
which cannot be fully retracted without potential global impact in the
rest of the exisiting reactors -- implying  a huge cost in
backprogation that would be problematic.

More generally tying a token
to future outcomes in a model that is meant to be confronted to rela
executionis something that is problematic. It may be valid ifg you can
gurantee that this will always happen (for example the reactor
managing \texttt{Agent.location} timeline guarantees that the outcome
of the \texttt{Going} will {\em always} result on ending at the traget
location) but should be manipulated while keeping in mind that
predicting the future is not absolute. In this case one could alter
this part of the model as follow in order to give the reactor more
flexibility:
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
The alteration made to relax the constraint on linbe 5 from a strict 
\texttt{equals} to contains and then add the new constraint at line 11
that ensure that the last \texttt{Going} done during this \texttt{Go}
ends up in the location the agent wnats to be. This firts level of
relaxation will therefore allow the agent to usies multiple
\texttt{Going} should the first one fail to reach our objective.  The
probalem can even be relaxed further if needed but this relaxation
gicves already more felxibility for the reactor to recover from local
failures.

>>>>

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition?} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \rx reactor execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

The manner in which the agent interleaves planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization occurs at every tick while
interrupting planning; this allows a reactor to identify its current
state that can propagate through the reactor hierarchy in order to
ensure a consistent view of the state of the world between all the
reactors.
% This view of a reactor has two parallel tasks allowing for long
% planning deliberation while ensuring that the state evolution
% propagates through the agent as time advance.
These intertwined tasks together provide two natural information flows
(bottom-up for state estimation and top-down for plan projection) that
we expect to see in a control loop. As they are both inference-based
processes it is natural to implement a reactor based on generic
automated planning. For example in Fig. \ref{fig:agent}, apart from
\textsf{Shore comm.} and \textsf{Vehicle} acting as interfaces to the
system, every other reactor (in blue) are different instances of a \eu
reactor differentiated only by their model and \eu solver
configurations.

@@

Within \rx, an instance of an \eu reactor comes with the sole purpose
to deliberate. A typical design of such a reactor takes its cues from
\texttt{IDEA} \cite{mus02, mus06} where both planning and execution
are based on a single unified model within a rich representation that
provides support for temporal inference. In \rx, the execution cycle
consists of two deliberation processes which are implemented as two
\eu solvers modifying the \emph{same} plan database. %  and for which the
% execution is managed by the \rx framework as time advance during the
% execution of the system:

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. \comment{Need
  to develop the reason why it is a necessity: primarily to provide
  its internal state to whoever observe it but also simply to ensure
  that its current representation of the world is up to date and still
  consistent}

We introduce this requirement as a new type of flaw (Section
\ref{sec:europa:arch}) for the \eu framework. This flaw forces the
reactor to identify its internal state for the current tick. By using
this flaw one can describe the synchronization process as:

@@

An internal state flaw can be resolved using one of the following
choices which are evaluated in order:

@@

\item Extend the previous state value to end after this tick (\ie
  restrict its end time to $[\tau+1, \infty)$). \comment{first use of
    $\tau$, so needs to be defined earlier}

\item Start the next active token in the timeline 
  by restricting its start time to the single value $\tau$

@@

All of these choices are evaluated sequentially until a consistent
solution with no flaws at this tick, is identified. In order to do so
we need to make the assumption that the current state value of an
internal timeline does not depend on the future or more accurately
that any choices made during this synchronization will not lead to a
future inconsistency. Such an assumption impacts the set of possible
domains the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model :

\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   equals(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11 }
\end{verbatim}

While this model is perfectly standard and acceptable in the overall
case. It can become problematic when tied to execution. Specifically it
puts a strong tie between the {\em fact} that we are \texttt{Going} to a
location and the {\em expected} future outcome of ending at this
location. This ties is even stronger due to the constraint at line 5.
Knowing this aspect assume that the \texttt{Agent.location} is {\em
  external} to the reactor which receive the observation
\texttt{Going(Home,SuperMarket)} which synchronization resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the Agent
timeline. From this point this means that the only possible outcome on
\texttt{Agent.location} is to be \texttt{At(SuperMarket)} in the
directly foreseeable future. In the real world the fact that we are
\texttt{Going} to a certain location does not necessarly imply that we
will succeed on ending up there. An extreme case could be that due to
some exogenous event, the road to the \texttt{SuperMarket} is closed
which could result on an alternate observation \texttt{At(7th st. \& Foo
  Ave.)}. This observation contradicts thre existing plan the reactor
maintin and it is important here to relaize the temp[oral direction of
the causal link that is triggereing this inconsitency. Indeed at the
time we will observe this contradicting \texttt{At}, the rule that
generated this conlict is connected to the past observations
\texttt{Going} and \texttt{Go} which cannot be fully retracted without
potential global impact in the rest of the exisiting reactors --
implying a huge cost in backprogation that would be problematic.

More generally tying a token to future outcomes in a model that is meant
to be confronted to rela 
executionis something that is problematic. It may be valid ifg you can
gurantee that this will always happen (for example the reactor
managing \texttt{Agent.location} timeline guarantees that the outcome
of the \texttt{Going} will {\em always} result on ending at the traget
location) but should be manipulated while keeping in mind that
predicting the future is not absolute. In this case one could alter
this part of the model as follow in order to give the reactor more
flexibility:

\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}

The alteration made to relax the constraint on linbe 5 from a strict 
\texttt{equals} to contains and then add the new constraint at line 11
that ensure that the last \texttt{Going} done during this \texttt{Go}
ends up in the location the agent wnats to be. This firts level of
relaxation will therefore allow the agent to usies multiple
\texttt{Going} should the first one fail to reach our objective.  The
probalem can even be relaxed further if needed but this relaxation
gicves already more felxibility for the reactor to recover from local
failures.
